A planetary collision afterglow and transit of the resultant debris cloud

### 2023-10-12 src/tex/ms.tex
[showyourwork] overleaf sync

How the emission from a post-impact body would evolve with time is highly dependent on the initial mass distribution and thermal state of the body, and the balance between radiative cooling, viscous spreading, and mass and angular momentum transport by condensates \cite{Lock2017,Lock18,Lock2020}. 
%
Given the limitations of SPH simulations (see above) it is not possible to accurately determine the initial structure of post-impact bodies in the relevant regime.
%
To explore a range of possible evolution pathways, we have calculated the evolution of post-impact bodies with different power-law surface density profiles, $\Sigma$, profiles, under the limiting case that radiative cooling and condensation of the vapor is the single driver for evolution of the structure. 
%
We chose a power-law surface density profile as it can straightforwardly cover the wide range of surface density profiles expected after super-Earth/mini-Neptune collisions, based on that found in impact simulations between lower-mass, terrestrial bodies \cite{Lock2017,Canup2001,Canup2012,Cuk2012,Rufu2017,Reufer2012}. 
%
The surface density profile is given by:
%
\begin{equation}
\Sigma=\begin{cases}
\Sigma_0, & \text{if } r_{xy}\leq R_{\rm c}\\
\alpha r_{xy}^{\beta}, & \text{if } R_{\rm c} > r_{xy}\leq R_{\rm emit}\\
0, & \text{if } r_{xy}> R_{\rm emit}
\end{cases}   
\label{eqn:sigma}
\end{equation}
%
where: $r_{xy}$ is the distance from the rotation axis; $R_{\rm c}$ is the outer radius of a constant surface density central region, roughly analogous to the corotating regions seen in Earth-mass synestias \cite{Lock2017,Lock18}; $\Sigma_0$ is the surface density of the central region; $\beta$ is the power law exponent; and $R_{\rm emit}$ is the initial emitting radius. Imposing surface density continuity gives
%
\begin{equation}
    \alpha=\Sigma_0 R_{\rm c}^{-\beta} \; .
\end{equation}
%
We can determine $\Sigma_0$ by fixing the mass of the body, 
%
\begin{equation}
    M_{\rm p} = \int_0^{R_{\rm c}}{2 \pi r'_{xy} \Sigma_0 dr'_{xy}} + \int_{R_{\rm c}}^{R_{\rm emit}}{2 \pi r'_{xy} \Sigma(r'_{xy}) dr'_{xy}} \; ,
\end{equation}
%
which can be solved to give:
%
\begin{equation}
\Sigma_0=\begin{cases}
\frac{M_{\rm p}}{\pi \left [R_{\rm c}^2 + 2 R_{\rm c}^{-\beta} \left ( \ln{R_{\rm emit}}-\ln{R_{\rm c}}\right ) \right ]}, & \text{if } \beta=-2,\\
\frac{M_{\rm p}}{\pi \left [R_{\rm c}^2 + \frac{2 R_{\rm c}^{-\beta}}{\beta+2} \left ( R_{\rm emit}^{\beta+2}-R_{\rm c}^{\beta+2}\right ) \right ]}, & \text{otherwise.} 
\end{cases}
\end{equation}
%
The time taken for a given region of the structure to cool to the point that a sufficient fraction of material is condensed for the temperature to drop below the condensation-buffer (see above) and so the emitted flux to drop is given by
%
\begin{equation}
    t_{\rm cool} (r_{xy})=\frac{l f(r_{xy}) \Sigma(r_{xy})}{\sigma T^4_{\rm emit}}
\end{equation}
%
where $f$ is the initial vapor fraction at that radius, $l$ is the latent heat of vaporization of the material (here we have taken the limiting case of pure water $l=2.256\times 10^6$ \cite{Chase1998}, but the addition of silicates could make the latent heat much larger), $\sigma$ is the Stefan–Boltzmann constant, and $T_{\rm emit}$ is the emission temperature.
%
Figure~\ref{fig:Hill_Bondi_R} shows example evolutions of the emission from a post-impact body using this model for different parameters (solid lines).
%
Results of a modified model where the initial surface density (by the addition of a constant parameter to Equation~\ref{eqn:sigma}) is forced to be zero at $R_{\rm emit}$ are shown as dashed lines.

>>>>

How the emission from a post-impact body would evolve with time is highly dependent on the initial mass distribution and thermal state of the body, and the balance between radiative cooling, viscous spreading, and mass and angular momentum transport by condensates \cite{Lock2017,Lock18,Lock2020}. 
%
Given the limitations of SPH simulations (see above) it is not possible to accurately determine the initial structure of post-impact bodies in the relevant regime.
%
To explore a range of possible evolution pathways, we have calculated the evolution of post-impact bodies with different power-law surface density profiles, $\Sigma$, under the limiting case that radiative cooling and condensation of the vapor is the single driver for evolution of the structure. 
%
We chose a power-law surface density profile as it can straightforwardly cover the wide range of surface density profiles expected after super-Earth/mini-Neptune collisions, based on that found in impact simulations between lower-mass, terrestrial bodies \cite{Lock2017,Canup2001,Canup2012,Cuk2012,Rufu2017,Reufer2012}. 
%
The surface density profile is given by:
%
\begin{equation}
\Sigma=\begin{cases}
\Sigma_0, & \text{if } r_{xy}\leq R_{\rm c}\\
\alpha r_{xy}^{\beta}, & \text{if } R_{\rm c} < r_{xy}\leq R_{\rm emit}\\
0, & \text{if } r_{xy}> R_{\rm emit}
\end{cases}   
\label{eqn:sigma}
\end{equation}
%
where: $r_{xy}$ is the distance from the rotation axis; $R_{\rm c}$ is the outer radius of a constant surface density central region, roughly analogous to the corotating regions seen in Earth-mass synestias \cite{Lock2017,Lock18}; $\Sigma_0$ is the surface density of the central region; $\beta$ is the power law exponent; and $R_{\rm emit}$ is the initial emitting radius. Imposing surface density continuity gives
%
\begin{equation}
    \alpha=\Sigma_0 R_{\rm c}^{-\beta} \; .
\end{equation}
%
We can determine $\Sigma_0$ by fixing the mass of the body, 
%
\begin{equation}
    M_{\rm p} = \int_0^{R_{\rm c}}{2 \pi r'_{xy} \Sigma_0 dr'_{xy}} + \int_{R_{\rm c}}^{R_{\rm emit}}{2 \pi r'_{xy} \Sigma(r'_{xy}) dr'_{xy}} \; ,
\end{equation}
%
which can be solved to give:
%
\begin{equation}
\Sigma_0=\begin{cases}
\frac{M_{\rm p}}{\pi \left [R_{\rm c}^2 + 2 R_{\rm c}^{-\beta} \left ( \ln{R_{\rm emit}}-\ln{R_{\rm c}}\right ) \right ]}, & \text{if } \beta=-2,\\
\frac{M_{\rm p}}{\pi \left [R_{\rm c}^2 + \frac{2 R_{\rm c}^{-\beta}}{\beta+2} \left ( R_{\rm emit}^{\beta+2}-R_{\rm c}^{\beta+2}\right ) \right ]}, & \text{otherwise.} 
\end{cases}
\end{equation}
%
The time taken for a given region of the structure to cool to the point that a sufficient fraction of material is condensed for the temperature to drop below the condensation-buffer (see above) and so the emitted flux to drop is given by
%
\begin{equation}
    t_{\rm cool} (r_{xy})=\frac{l f(r_{xy}) \Sigma(r_{xy})}{\sigma T^4_{\rm emit}}
\end{equation}
%
where $f$ is the initial vapor fraction at that radius, $l$ is the latent heat of vaporization of the material (here we have taken the limiting case of pure water $l=2.256\times 10^6$ \cite{Chase1998}, but the addition of silicates could make the latent heat much larger), $\sigma$ is the Stefan–Boltzmann constant, and $T_{\rm emit}$ is the emission temperature.
%
Figure~\ref{fig:Hill_Bondi_R} shows example evolutions of the emission from a post-impact body using this model for different parameters (solid lines).
%
Results of a modified model where the initial surface density (by the addition of a constant parameter to Equation~\ref{eqn:sigma}) is forced to be zero at $R_{\rm emit}$ are shown as dashed lines.
